Embedded Scalars, Preferred Readings \& Intonation: {A}n Experimental Revisit

### 2014-08-29 text/paper/paper4.tex
integrated comments by Fabian and Petra

\begin{abstract}
  Scalar item \emph{some} is widely assumed to receive a pragmatic
  meaning enrichment to \emph{some but not all} if it occurs in matrix
  position. The question under which circumstances this enrichment can
  occur in certain embedded positions plays an important role in
  deciding how to delineate semantics and pragmatics. % Some claim
  % that it requires marked intonation for the relevant pragmatic
  % enrichment to occur in embedded position; others allow for it more
  % freely.
  We present new experimental data that bears on this theoretical
  issue. In distinction to previous experimental approaches, we
  presented sentence material auditorily in order to explicitly
  control prosodic markedness of the scalar item. Moreover, our
  experimental design gives insight into the preference of
  hypothesized meaning differences, not only on their general
  availability. The data obtained in this way turns out problematic
  for both a traditional Gricean view of scalar implicature
  calculation, as well as the more recent grammatical approach.
\end{abstract}

@@

The existential quantifier \emph{some} is usually assumed to receive a
semantic interpretation similar to logical $\exists$, so that
(\ref{bsp:Plain-SI-Target}) is literally true even when Hans solved
all of the problems. But it is also usually considered to invite
comparison with (at least) the semantically stronger universal
quantifier \emph{all}
\citep[c.f.][]{Horn1972:On-the-Semantic,Gazdar1979:Pragmatics:-Imp,AtlasLevinson1981}. This
scalar comparison can lead to an upper-bound meaning enrichment, e.g.,
when an utterance of (\ref{bsp:Plain-SI-Target}) is taken to invite
the inference in (\ref{bsp:Plain-SI-Implicature}).

@@

\noindent The classical explanation of this inference, following the
pioneering work of \citet{Grice1975:Logic-and-Conve}, is that
(\ref{bsp:Plain-SI-Implicature}) is a pragmatic inference, a so-called
\emph{quantity implicature}, derived by an abductive inference as the
best explanation of why informed, knowledgable and cooperative
speakers would utter (\ref{bsp:Plain-SI-Target}) when they could also
utter the semantically stronger and relevant
(\ref{bsp:Plain-SI-Alternative}) \citep[see][for
recent overview]{Geurts2010:Quantity-Implic}.

@@

This paper deals with the interpretation of two types of sentences,
where the scalar item \emph{some} occurs in the scope of other logical
operators. In \as-sentences (mnemonic for \textit{all \dots some
  \dots}) as in (\ref{bsp:AE}) the scalar item \emph{some} is embedded
under universal quantifier \emph{all}. In \es-sentences (mnemonic for
\textit{exactly one \dots some \dots}) like (\ref{bsp:GE}) \emph{some}
takes scope under the non-monotonic quantifier \emph{exactly one}.
According to current pragmatic theory, there are at least three
relevant candidate readings for \as- and \es-sentences: (i) a
\emph{literal reading} like in (\ref{bsp:AE-Literal}) and
(\ref{bsp:GE-Literal}) where \emph{some} has only its literal meaning;
(ii) a \emph{global reading} like in (\ref{bsp:AE-Global}) and
(\ref{bsp:GE-Global}) where we enrich the meaning of (\ref{bsp:AE})
and (\ref{bsp:GE}) with the negation of the corresponding sentences
(\ref{bsp:AE-Alternative}) and (\ref{bsp:GE-Alternative}) where
\emph{some} is replaced by \emph{all}; and also (iii) a \emph{local
  reading} like in (\ref{bsp:AE-Local}) and (\ref{bsp:GE-Local}) where
\emph{some} is interpreted as \emph{some but not all} in the scope of
the embedding quantifier.

@@

\noindent Given this plurality of hypothesized readings, we need to
ask:
\begin{enumerate}[Q1:]
\item which of these readings are available to na\"{i}ve subjects; and
\item which of the available readings are preferred.\fn{It may
    ultimately be impossible to keep unavailability and strong
    dispreference strictly apart. For our purposes this is not
    important. Every time we speak of availability, the so-inclined
    reader may think of non-negligible preference.}
\end{enumerate}
Addressing these questions empirically is relevant because they lie at
the heart of the current debate about the exact location and nature of
the interface between semantics and pragmatics. The controversy
concerns the question under which circumstances and to what extent the
upper-bound inference from \emph{some} to \emph{some but not all} can
occur within the compositional computation of semantic values. Two
opposing theoretical positions can be distinguished here:
\emph{traditionalism} holds in particular that local readings should
not be readily attested unless marked by special intonation;
\emph{grammaticalism} holds that local readings are available and, at
least under the most prominent of selection principles for candidate
readings, even preferred (see Section~\ref{sec:grammaticalism}). To
distinguish between these opposing positions, it is therefore
necessary to assess the availability and relative preference of
readings for \as- and \es-sentences under conditions that control for
intonational markedness. This is what this paper aims to do.

A number of empirical studies have already studied sentences like
(\ref{bsp:AE}) and (\ref{bsp:GE})
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. Unfortunately,
results from these studies do not provide uncontroversial
evidence. This is due to several reasons. Firstly, previous studies
have only accumulated limited evidence pertaining to the preference
relations between candidate readings of both types of
sentences. Secondly, as \citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Scalar-expressi} argue, the alleged
evidence for the availability of local readings of \as-sentences
suggested by studies of \citet{CliftonDube2010:Embedded-Implic} and
\citet{ChemlaSpector2010:Experimental-Ev}, might well be confounds
with what van Tiel and Geurts call \emph{typicality effects}. Finally,
and most importantly, previous studies presented target sentences
visually. This is highly problematic, because traditionalists admit to
the availability of local readings as long as it is intonationally
marked
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Scalar-expressi}. We
will address this extant position as the \emph{intonational markedness
  hypothesis}. But if intonation plays a role, it is vital to present
target sentences auditorily, so as to be able to control for effects
of ``silent prosody'' that subjects may apply when reading a sentence
\citep[see][]{Bader98,Fodor98}.

This paper therefore presents results form an experimental study
designed to deal with the above mentioned problems. We presented
sentence material auditorily and explicitly manipulated intonational
stress on embedded scalar items. To cater for potential pictorial
confounds, such as from typicality, we employed an \emph{incremental
  verification task} in which parts of a to-be-evaluated picture were
hidden and only revealed piece-wise at the participants' request until
they felt able to give a binary truth-value judgement
\citep[see][]{Conroy2008}. This way we obtained categorical behavioral
data that is both clearly indicative of the reading that subjects
assumed and fairly independent of any potential pictorial
confounds. Moreover, this method also gives information about reading
preferences. We validated this by including a control condition with
ambiguous sentences with a well-known preference on available
readings.

The data obtained by this method is problematic for both
traditionalism and grammaticalism. Contrary to the predictions of
traditionalism, we observed many answers indicative of local readings
of \es-sentences, irrespective of intonational markedness. But also
grammaticalism is incongruent with the observed pattern in our
data. Backed up by the strongest meaning hypothesis
\citep{DalrympleKanazawa1998:Reciprocal-Expr}, grammaticalism predicts
local readings to be the preferred readings, but we found no evidence
for that. More strongly even, the preference order among putative
readings suggested by our data cannot even in principle be explained
by comparisons of logical strength among putative readings. In sum,
our data sheds doubt on both traditionalism and grammaticalism in
their currently most common forms. We conclude the paper by some
speculation about how either position might be amended in the light of
our data.

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out the
predictions of traditionalism and grammaticalism about availability
and preference of readings for these
sentences. Section~\ref{sec:previous-studies} reviews related
experimental studies to the extent necessary to motivate our own
approach. Section~\ref{sec:exp} eventually contributes our
experimental design and our results.

@@

An \as-sentence like (\ref{bsp:AE}), repeated below, has a literal
reading (\lit) as in (\ref{bsp:AE-Literal}), a global reading (\glb) as in
(\ref{bsp:AE-Global}) and a local reading (\loc) as in (\ref{bsp:AE-Local}).

@@

Given these entailment relations, there are four kinds of situations,
the names for which we borrow from
\citet{ChemlaSpector2010:Experimental-Ev}, that we can distinguish
based on different truth values for our candidate readings. These are
given in Table~\ref{fig:situations-AE}.
%
\begin{table}
  \centering
  \subfloat[][\as-sentences]{
  \label{fig:situations-AE}
    \begin{tabular}{lccc}
    \toprule
    situation    & \multicolumn{3}{c}{reading} 
  \\ 
  \cmidrule(r){2-4}
     & \lit & \glb & \loc \\ \midrule
    false   & 0 & 0 & 0 \\
    literal & 1 & 0 & 0 \\
    weak    & 1 & 1 & 0 \\
    strong  & 1 & 1 & 1 \\ \bottomrule
  \end{tabular}
} \qquad \qquad
  \subfloat[][\es-sentences]{
  \label{fig:situations-GS}
  \begin{tabular}{lccc}
    \toprule
    situation    & \multicolumn{3}{c}{reading} 
  \\ 
  \cmidrule(r){2-4}
    & \textsc{lit} & \textsc{glb} & \textsc{loc} \\ \midrule
    false   & 0 & 0 & 0 \\
    literal & 1 & 0 & 0 \\
    local   & 0 & 0 & 1 \\
    all     & 1 & 1 & 1 \\ \bottomrule
  \end{tabular}
}
\caption{Possible truth-value distributions for readings of \as- and \es-sentences}
\label{tab:situations}
\end{table}
%
Examples of these kinds of situations are given in
Figure~\ref{fig:AS-distinguishing-pics}, where the dots on the right
of each diagram represent students, the dots on the left represent
papers and an arrow from a student to a paper indicates that the
student read the paper. Notice that other arrangements of arrows might
equally well serve as examples for the various situations.

@@

  \caption{Distinguishing scenarios for \as-sentences}
  \label{fig:AS-distinguishing-pics}
\end{figure}

@@

The situation for \es-sentences like (\ref{bsp:GE}) is similar but a
little more complicated because the embedding quantifier is
non-monotonic. Again, we consider a literal reading as in
(\ref{bsp:GE-Literal}), repeated below, a global reading as in
(\ref{bsp:GE-Global}) and a local reading as in (\ref{bsp:GE-Local}).

@@

\noindent Entailment relations in this case are non-linear:
\begin{exe}
  \ex \label{bsp:Entailments-GE}       \lit\ $\supset$ \glb\ $\color{mycol}{\subset}$ \loc 
    % \hspace{1cm} and \hspace{1cm}  \lit\ $\not \supset$ \loc
\end{exe}
By definition of global readings $\glb$ entails $\lit$. This
entailment is asymmetric because the extra information that it is not
the case that exactly one student read all of the papers is not
entailed by the literal reading (\ref{bsp:GE-Literal}). However,
unlike for \as-sentences, \loc is not stronger than \glb, but
asymmetrically entailed by the latter. To see this, notice that the
global reading is equivalent to:

@@

\noindent Finally, \loc and \lit are logically independent: all
possible combinations of truth-values for \loc and \lit are possible.

Given these entailment relations, there are again four different
situations corresponding to the four possible distributions of truth
values for candidate readings. These are given in
Table~\ref{fig:situations-GS} and named following
\citet{ChemlaSpector2010:Experimental-Ev}. Examples for each situation
are given in Figure~\ref{fig:ES-distinguishing-pics}.

@@

  \caption{Distinguishing scenarios for \es-sentences}
  \label{fig:ES-distinguishing-pics}

@@

There are two main theoretical positions which make different
predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation}. We
will refer to these here as \mymark{traditionalism} and
\mymark{grammaticalism} and treat each one in turn. Since there is
some leeway in assessing the predictions for each (depending on which
of several additional assumptions to adopt), we will distinguish
different varieties of both. Predictions of all relevant variants are
summarized in Table~\ref{tab:predictions} on
page~\pageref{tab:predictions}.

Although strictly speaking not a part of a Gricean theory of scalar
implicature calculation, traditionalists frequently concede that local
readings are in principle available, albeit not as a standard scalar
implicature, but as a different pragmatic process triggered by
prosodic markedness. We examine this \emph{intonational markedness
  hypothesis} separately in Section~\ref{sec:inton-mark-hypoth},
alongside a brief revision of the relevant experimental data.

@@

The Gricean recipe applies also to \as- and \es-sentences and derives
the global reading in a straightforward way. Consequently,
traditionalism predicts that both literal and global readings are
available: literal readings, because these form the starting point of
pragmatic reasoning; global readings because these may be arrived at
by the Gricean recipe.

Traditionalism's predictions about preferences between literal and
global readings depend on whether the auxiliary assumptions necessary
to derive global readings by the Gricean recipe are plausibly
met. These auxiliary assumptions include relevance of the extra
information provided by the global reading, mutual awareness that the
stronger alternative has been a speaker option, the speaker's
competence about the issue, etc. Normally, traditionalist accounts
would assume that these extra assumptions are met. In that case,
traditionalism would predict that global readings should be preferred
over literal readings. On the other hand, if auxiliary assumptions are
implausible, traditionalism would predict that the literal readings
would be preferred over the global one. Consequently, there are at
least two varieties of traditionalism that, depending on which
additional assumptions we would make, yield slightly different
predictions: \mymark{the strong variety of traditionalism} maintains
that the auxiliary assumptions of the Gricean recipe hold
usually/strongly/unless-completely-intenable and therefore predicts
that global readings are preferred over literal ones; \mymark{the weak
  variety of traditionalism} holds that the auxiliary assumptions are
more fragile and predicts that literal readings are preferred over the
global ones.\fn{We mention for clarity that although weak
  traditionalism does not predict the global reading to be preferred,
  it might still predict an \emph{epistemically weak implicature},
  similar to the global reading, that the speaker is uncertain whether
  the stronger alternative is true. (See also the discussion in
  Section~\ref{sec:discussion-1})}

@@

\noindent Clearly, if we conjoin a literal reading of (\ref{bsp:AE})
with the negation of (\ref{bsp:AE-Alternatives-Extended}) we obtain
exactly the local reading. (Notice that the negation of
(\ref{bsp:AE-Alternatives-Extended}) entails the negation of
(\ref{bsp:AE-Alternative}), so it makes no difference (not) to add it
in this case.) Consequently, traditionalism does predict that the
local reading of \as-sentences is available if
(\ref{bsp:AE-Alternatives-Extended}) is an available alternative. We
should therefore introduce another distinction: \mymark{restricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
not available and therefore predicts that the local reading for
\as-sentences is not available; in contrast, \mymark{unrestricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
available and so is the local reading.

Still, even unrestricted traditionalism would not predict that the
local reading is preferred over the global reading. This is because
the derivation of the local reading via
(\ref{bsp:AE-Alternatives-Extended}) hinges on yet another auxiliary
assumption, namely the availability of the alternative in
(\ref{bsp:AE-Alternatives-Extended}), which seems much less obvious
and therefore presumably is less readily available than the
alternative in (\ref{bsp:AE-Alternative}). Consequently, we derive the
following predictions about preferences for unrestricted
traditionalism: weak unrestricted traditionalism prefers the literal
over the global over the local reading, while strong unrestricted
traditionalism prefers the global over the local over the literal
reading (see also Table~\ref{tab:predictions}).

@@

Recently, a lot of evidence based on intuitive judgements and more
theoretical considerations has been discussed in favor of
\mymark{grammaticalism}
\citep[see][]{Chierchia2006:Broaden-Your-Vi,Fox2007:Free-Choice-and,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical}. % When
% it comes to the predictions of available readings of \as- and
% \es-sentences, grammaticalism appears like the conjunction of
% traditionalism and conventionalism. But conceptually speaking,
% grammaticalism goes down a road of its own.
Grammaticalism approaches quantity implicatures by postulating a
silent operator that can be variably applied during compositional
computation of a sentence's truth-value
\citep{Chierchia2006:Broaden-Your-Vi}, if necessary multiple times
\citep{Fox2007:Free-Choice-and}. This silent operator is referred to
as $\mathrm{O}(\cdot)$ or $\exh(\cdot)$, because it is assumed to be
similar in effect to the meaning of particle \emph{only} or of the
mechanism of \emph{exhaustive interpretation}
\citep{GroenendijkStokhofThesis1984,Stechowvon-StechowZimmermann1984:Term-Answers-an,Rooijvan-RooijSchulz2013:Exhaustive-Inte,vanRooijSchulz:ExhaustiveInterpretation,Fox2007:Free-Choice-and}. For
our purposes, it is enough to note that $\exh(\cdot)$ is a poly-typed
function that enriches an expression $X$, which need not be a full
proposition, based on a set $\alt(X)$ of (suitable, relevant)
alternatives to $X$ that yields an enriched meaning of the
form:\fn{More sophisticated formulations of exhaustive interpretation
  have been proposed
  \citep[e.g.][]{Schulz2005:A-Pragmatic-Sol,vanRooijSchulz:ExhaustiveInterpretation,Spector2006:Scalar-Implicat,Fox2007:Free-Choice-and}
  but this simple formulation is sufficient for the purposes of this
  paper.}
\begin{exe}
  \ex \label{bsp:Exh-Def} $\exh(X,\alt(X)) = X \bigwedge_{A \in
      \alt(X)} \neg A$.
\end{exe}

>>>>

\begin{abstract}
  The scalar item \emph{some} is widely assumed to receive a pragmatic
  meaning enrichment to \emph{some but not all} if it occurs in matrix
  position. The question under which circumstances this enrichment can
  occur in certain embedded positions plays an important role in
  deciding how to delineate semantics and pragmatics. % Some claim
  % that it requires marked intonation for the relevant pragmatic
  % enrichment to occur in embedded position; others allow for it more
  % freely.
  We present new experimental data that bear on this theoretical
  issue. In distinction to previous experimental approaches, we
  presented sentence material auditorily in order to explicitly
  control prosodic markedness of the scalar item. Moreover, our
  experimental sheds light on the relative preferences or salience of
  candidate readings. The presented data turns out to be problematic
  for both a traditional Gricean view of scalar implicature
  calculation, as well as the more recent grammatical approach.
\end{abstract}

@@

The existential quantifier \emph{some} is usually assumed to receive a
semantic interpretation similar to logical $\exists$, so that
(\ref{bsp:Plain-SI-Target}) is literally true even when Hans solved
all of the problems. But use of \emph{some} is also usually considered
to invite comparison with (at least) the semantically stronger
universal quantifier \emph{all}
\citep[c.f.][]{Horn1972:On-the-Semantic,Gazdar1979:Pragmatics:-Imp,AtlasLevinson1981}. This
scalar comparison can lead to an upper-bound meaning enrichment, e.g.,
when an utterance of (\ref{bsp:Plain-SI-Target}) is taken to invite
the inference in (\ref{bsp:Plain-SI-Implicature}).

@@

\noindent The classical explanation of this inference, following the
pioneering work of \citet{Grice1975:Logic-and-Conve}, is that
(\ref{bsp:Plain-SI-Implicature}) is a pragmatic inference, a so-called
\emph{quantity implicature}, derived by an abductive inference as the
best explanation of why informed, knowledgable and cooperative
speakers would utter (\ref{bsp:Plain-SI-Target}) when they could also
utter the semantically stronger and relevant
(\ref{bsp:Plain-SI-Alternative}) \citep[see][for a recent
overview]{Geurts2010:Quantity-Implic}.

@@

Whereas the existence of such implicatures is rather uncontroversial
for plain occurrences of \emph{some}, it is still an issue of debate
whether comparable enrichments are also found in so-called
``embedded'' cases, where the scalar item \emph{some} occurs in the
scope of other logical operators. This paper deals with to such cases.
In \as-sentences (mnemonic for \textit{all \dots some \dots}) as in
(\ref{bsp:AE}) the scalar item \emph{some} is embedded under universal
quantifier \emph{all}. In \es-sentences (mnemonic for \textit{exactly
  one \dots some \dots}) like (\ref{bsp:GE}) \emph{some} takes scope
under the non-monotonic quantifier \emph{exactly one}.  According to
current pragmatic theory, there are at least three relevant candidate
readings for \as- and \es-sentences: (i) a \emph{literal reading} like
in (\ref{bsp:AE-Literal}) and (\ref{bsp:GE-Literal}) where \emph{some}
has only its literal meaning; (ii) a \emph{global reading} like in
(\ref{bsp:AE-Global}) and (\ref{bsp:GE-Global}) where the meaning of
(\ref{bsp:AE}) is enriched and (\ref{bsp:GE}) with the negation of the
corresponding sentences (\ref{bsp:AE-Alternative}) and
(\ref{bsp:GE-Alternative}) respectively; and also (iii) a \emph{local
  reading} like in (\ref{bsp:AE-Local}) and (\ref{bsp:GE-Local}) where
\emph{some} is interpreted as \emph{some but not all} in the scope of
the embedding quantifier.

@@

\noindent Given this plurality of hypothesized readings, we need to
ask:
\begin{enumerate}[Q1:]
\item which of these readings are available to na\"{i}ve subjects; and
\item which of the available readings are preferred.\fn{It may
    ultimately be impossible to keep unavailability and strong
    dispreference strictly apart. For our purposes this is not
    important. Every time we speak of availability, the so-inclined
    reader may think of non-negligible preference.}
\end{enumerate}
Addressing these questions empirically is relevant because they lie at
the heart of a current debate about the nature of quantity inferences
and the demarkation between semantics and pragmatics. The controversy
concerns the question under which circumstances and to what extent the
upper-bound inference from \emph{some} to \emph{some but not all} can
occur within the compositional computation of semantic values. Two
main theoretical positions can be distinguished here:
\emph{traditionalism} holds in particular that local readings should
not occur frequently unless marked by special intonation;
\emph{grammaticalism} holds that local readings are available and, at
least under the most prominent of selection principles for candidate
readings, even preferred (see Section~\ref{sec:grammaticalism}). To
distinguish between these opposing positions, it is necessary to
assess the availability and relative preference of readings for \as-
and \es-sentences under conditions that control for prosodic
markedness.

A number of empirical studies have already examined sentences like
(\ref{bsp:AE}) and (\ref{bsp:GE})
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. Unfortunately,
results from these studies do not provide uncontroversial
evidence. This is due to several reasons. Firstly, previous studies
have only accumulated limited evidence pertaining to the preference
relations between candidate readings of both types of
sentences. Secondly, as \citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue, the alleged
evidence for the availability of local readings of \as-sentences
suggested by studies of \citet{CliftonDube2010:Embedded-Implic} and
\citet{ChemlaSpector2010:Experimental-Ev}, might well be confounds of
what van Tiel and Geurts call \emph{typicality effects} (see
Section~\ref{sec:previous-studies} for discussion). Finally, and most
importantly, previous studies presented target sentences
visually. However, traditionalists admit to the availability of local
readings as long as the utterance is prosodically marked
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}. We
will address this extant position as the \emph{prosodic markedness
  hypothesis}. If intonation plays a role, it is vital to present
target sentences auditorily, and to control for effects of ``silent
prosody'' that subjects may apply when reading a sentence
\citep[see][]{Bader98,Fodor98}.

This paper presents results form an experimental study designed to
deal with the above mentioned problems. We presented sentence
materials auditorily and explicitly manipulated intonational stress on
embedded scalar items. To cater for potential pictorial confounds,
such as from typicality, we employed an \emph{incremental verification
  task} in which parts of a to-be-evaluated picture were hidden and
only revealed piece-wise at the participants' request until they felt
able to give a binary truth-value judgement
\citep[see][]{Conroy2008}. This way we obtained categorical behavioral
data that is both more clearly indicative of the reading that subjects
assumed and mor independent of any potential pictorial confounds, such
as the mentioned typicality effects. Moreover, this method also gives
information about reading preferences. We validated this by including
control conditions with ambiguous sentences with a well-known
preference on available readings.

% The data obtained by this method is problematic for both
% traditionalism and grammaticalism. Contrary to the predictions of
% traditionalism, we observed many answers indicative of local readings
% of \es-sentences, irrespective of intonational markedness. But also
% grammaticalism is incongruent with the observed pattern in our
% data. Backed up by the strongest meaning hypothesis
% \citep{DalrympleKanazawa1998:Reciprocal-Expr}, grammaticalism predicts
% local readings to be the preferred readings, but we found no evidence
% for that. More strongly even, the preference order among putative
% readings suggested by our data cannot even in principle be explained
% by comparisons of logical strength among putative readings. In sum,
% our data sheds doubt on both traditionalism and grammaticalism in
% their currently most common forms. We conclude the paper by some
% speculation about how either position might be amended in the light of
% our data.

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out the
predictions of traditionalism and grammaticalism about availability
and preference of readings for these
sentences. Section~\ref{sec:previous-studies} reviews related
experimental studies to the extent necessary to motivate our own
approach. Section~\ref{sec:exp} eventually contributes our
experimental design, analyses and results.

@@

An \as-sentence like (\ref{bsp:AE}), repeated below, has a literal
reading (\lit) as in (\ref{bsp:AE-Literal}), a global reading (\glb)
as paraphrased in (\ref{bsp:AE-Global}) and a local reading (\loc) as
in (\ref{bsp:AE-Local}).

@@

Given these entailment relations, there are four kinds of situations,
the names for which we borrow from
\citet{ChemlaSpector2010:Experimental-Ev}, that we can distinguish
based on different truth values for our candidate readings. These are
given in Table~\ref{fig:situations-AE}.
%
\begin{table}
  \centering
  \subfloat[][\as-sentences]{
  \label{fig:situations-AE}
    \begin{tabular}{lccc}
    \toprule
    situation    & \multicolumn{3}{c}{reading} 
  \\ 
  \cmidrule(r){2-4}
     & \lit & \glb & \loc \\ \midrule
    false   & 0 & 0 & 0 \\
    literal & 1 & 0 & 0 \\
    weak    & 1 & 1 & 0 \\
    strong  & 1 & 1 & 1 \\ \bottomrule
  \end{tabular}
} \qquad \qquad
  \subfloat[][\es-sentences]{
  \label{fig:situations-GS}
  \begin{tabular}{lccc}
    \toprule
    situation    & \multicolumn{3}{c}{reading} 
  \\ 
  \cmidrule(r){2-4}
    & \textsc{lit} & \textsc{glb} & \textsc{loc} \\ \midrule
    false   & 0 & 0 & 0 \\
    literal & 1 & 0 & 0 \\
    local   & 0 & 0 & 1 \\
    all     & 1 & 1 & 1 \\ \bottomrule
  \end{tabular}
}
\caption{Possible truth-value distributions for readings of \as- and \es-sentences}
\label{tab:situations}
\end{table}
%
Examples of these kinds of situations are shown in
Figure~\ref{fig:AS-distinguishing-pics}, where the dots on the right
of each diagram represent students, the dots on the left represent
papers and an arrow from a student to a paper indicates that the
student read the paper. % Notice that other arrangements of arrows might
% equally well serve as examples for the various situations.

@@

\caption{Examples of distinguishing situations for \as-sentences}
  \label{fig:AS-distinguishing-pics}
\end{figure}

@@

The case for \es-sentences like (\ref{bsp:GE}) is similar but a
little more complicated because the embedding quantifier is
non-monotonic. Again, we consider a literal reading as in
(\ref{bsp:GE-Literal}), repeated below, a global reading as in
(\ref{bsp:GE-Global}) and a local reading as in (\ref{bsp:GE-Local}).

@@

\noindent Entailment relations in this case are non-linearly ordered:
\begin{exe}
  \ex \label{bsp:Entailments-GE}       \lit\ $\supset$ \glb\ $\color{mycol}{\subset}$ \loc 
    % \hspace{1cm} and \hspace{1cm}  \lit\ $\not \supset$ \loc
\end{exe}
By definition of global readings $\glb$ entails $\lit$. This
entailment is asymmetric because the extra information that it is not
the case that exactly one student read all of the papers is not
entailed by the literal reading (\ref{bsp:GE-Literal}). However,
unlike for \as-sentences, \loc is not stronger than \glb, but
asymmetrically entailed by the latter. To see this, notice that the
global reading is equivalent to:

@@

\noindent Finally, \loc and \lit are logically independent: all
combinations of truth-values for \loc and \lit are possible.

Given these entailment relations, there are again four different
situations corresponding to the four possible distributions of truth
values for candidate readings. These are given in
Table~\ref{fig:situations-GS} on page~\pageref{fig:situations-GS} and
named following \citet{ChemlaSpector2010:Experimental-Ev}. Examples
for each situation are given in
Figure~\ref{fig:ES-distinguishing-pics}.

@@

  \caption{Examples of distinguishing situations for \es-sentences}
  \label{fig:ES-distinguishing-pics}

@@

There are two main theoretical positions which make different
predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation}. We
will refer to these as \mymark{traditionalism} and
\mymark{grammaticalism} and treat each one in turn. Since there is
some leeway in assessing the predictions for each approach (depending
on which of several additional assumptions to adopt), we will
distinguish different varieties of both of them. Predictions of all
relevant variants are summarized in Table~\ref{tab:predictions} on
page~\pageref{tab:predictions}.

Although strictly speaking not a part of a Gricean theory of scalar
implicature calculation, traditionalists frequently concede that local
readings are in principle available, albeit not as a standard scalar
implicature, but as a different pragmatic process triggered by
prosodic markedness. We address this \emph{prosodic markedness
  hypothesis} separately in Section~\ref{sec:inton-mark-hypoth},
alongside a brief revision of the relevant experimental data.

@@

As for \as- and \es-sentences, the Gricean recipe derives global
readings in a straightforward way. Consequently, traditionalism
predicts that both literal and global readings are available: literal
readings, because these form the starting point of pragmatic
reasoning; global readings because these may be arrived at by the
Gricean recipe.

Traditionalism's predictions about preferences between literal and
global readings depend on whether the auxiliary assumptions necessary
to derive global readings by the Gricean recipe are plausibly
met. These auxiliary assumptions include relevance of the extra
information provided by the global reading, mutual awareness that the
stronger alternative has been a speaker option, the speaker's
competence about the issue, etc. Normally, traditionalist accounts
would assume that these extra assumptions are met. In that case,
traditionalism would predict that global readings should be preferred
over literal readings. On the other hand, if auxiliary assumptions are
implausible, traditionalism would predict that literal readings would
be preferred over global ones. Consequently, there are at least two
varieties of traditionalism that, depending on which additional
assumptions we would make, yield slightly different predictions:
\mymark{the strong variety of traditionalism} maintains that the
auxiliary assumptions of the Gricean recipe hold (e.g., as a matter of
strong default) and therefore predicts that global readings are
preferred over literal ones; \mymark{the weak variety of
  traditionalism} holds that the auxiliary assumptions are more
fragile and predicts that literal readings are preferred over global
ones.\fn{We mention for clarity that although weak traditionalism does
  not predict global readings to be preferred, it might still
  predict an \emph{epistemically weak implicature}, similar to the
  global reading, that the speaker is uncertain whether the stronger
  alternative is true. (See also the discussion in
  Section~\ref{sec:discussion-1})}

@@

\noindent Clearly, if we conjoin a literal reading of (\ref{bsp:AE})
with the negation of (\ref{bsp:AE-Alternatives-Extended}) we obtain
exactly the local reading. (Notice that the negation of
(\ref{bsp:AE-Alternatives-Extended}) entails the negation of
(\ref{bsp:AE-Alternative}), so it makes no difference (not) to add it
in this case.) Consequently, traditionalism does predict that the
local reading of \as-sentences is available if
(\ref{bsp:AE-Alternatives-Extended}) is an available alternative. We
therefore introduce another distinction: \mymark{restricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
not available and therefore predicts that the local reading for
\as-sentences is not available; in contrast, \mymark{unrestricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
available and so is the local reading.

As far as preferences are concerned, even unrestricted traditionalism
would not predict that the local reading is preferred over the global
reading. This is because the derivation of the local reading via
(\ref{bsp:AE-Alternatives-Extended}) hinges on yet another auxiliary
assumption, namely the availability of the alternative in
(\ref{bsp:AE-Alternatives-Extended}), which seems much less obvious
and therefore presumably is less readily available than the
alternative in (\ref{bsp:AE-Alternative}). Consequently, we derive the
following predictions about preferences for unrestricted
traditionalism: weak unrestricted traditionalism prefers the literal
over the global over the local reading, while strong unrestricted
traditionalism prefers the global over the local over the literal
reading (see also Table~\ref{tab:predictions}).

@@

Recently, a lot of evidence based on intuitive judgements and more
theoretical considerations has been discussed in favor of
\mymark{grammaticalism}
\citep[see][]{Chierchia2006:Broaden-Your-Vi,Fox2007:Free-Choice-and,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical}. % When
% it comes to the predictions of available readings of \as- and
% \es-sentences, grammaticalism appears like the conjunction of
% traditionalism and conventionalism. But conceptually speaking,
% grammaticalism goes down a road of its own.
Grammaticalism approaches quantity implicatures by postulating a
silent operator that can be variably applied during compositional
computation of a sentence's truth-value
\citep{Chierchia2006:Broaden-Your-Vi}, if necessary multiple times
\citep{Fox2007:Free-Choice-and}. This silent operator is referred to
as $\mathrm{O}(\cdot)$ or $\exh(\cdot)$, because it is assumed to be
similar in effect to the meaning of particle \emph{only} or of the
mechanism of \emph{exhaustive interpretation}
\citep{GroenendijkStokhofThesis1984,Stechowvon-StechowZimmermann1984:Term-Answers-an,Rooijvan-RooijSchulz2013:Exhaustive-Inte,vanRooijSchulz:ExhaustiveInterpretation,Fox2007:Free-Choice-and}. For
our purposes, it is enough to note that $\exh(\cdot)$ is a poly-typed
function that enriches an expression $X$, which need not be a full
proposition, based on a set $\alt(X)$ of (suitable, relevant)
alternatives to $X$ that yields an enriched meaning of the
form:\fn{More sophisticated formulations of exhaustive interpretation
  have been proposed
  \citep[e.g.][]{Schulz2005:A-Pragmatic-Sol,vanRooijSchulz:ExhaustiveInterpretation,Spector2006:Scalar-Implicat,Fox2007:Free-Choice-and}.}
\begin{exe}
  \ex \label{bsp:Exh-Def} $\exh(X,\alt(X)) = X \bigwedge_{A \in
      \alt(X)} \neg A$.
\end{exe}
